%% LaTeX2e class for student theses
%% sections/conclusion.tex
%% 
%% Karlsruhe Institute of Technology
%% Institute for Program Structures and Data Organization
%% Chair for Software Design and Quality (SDQ)
%%
%% Dr.-Ing. Erik Burger
%% burger@kit.edu
%%
%% Version 1.3.5, 2020-06-26

\chapter{Conclusion}
\label{ch:Conclusion}

In this research we determined which DLT characteristics are suitable for personal DP through mapping the characteristics of Hyperledger Fabric and Ethereum to fundamental personal DP requirements, as well as specific use case requirements in the field of healthcare and finance. While both use cases share the fundamental, as well as other secondary DP requirements, healthcare seems to be focused more on patient engagement, responsibility and availability in emergency situations, whereas finance places higher importance on disintermediation, transaction speed, cost, corruption and crime reduction.

We concluded that neither of our considered approaches is necessarily better than the other and they both exhibit characteristics which are beneficial in different situations. DLTs in general exhibit characteristics such as \textit{Turing-complete Smart Contracts, Traceability, Integrity}, etc., which make them highly suitable for personal DP, however, permissioned designs remain a preference, which can be largely motivated by the improvements in \textit{Throughput, Scalability, Performance} and enhanced \textit{Confidentiality} due to the restricted access to the ledger. On the other hand, public designs' advantages include higher level of \textit{Decentralisation} and \textbf{Trust}, which align with DLT's original philosophy.

However, the need for \textbf{Interoperability}, \textbf{Scalability} and \textbf{Compliance} in general remains as one of the biggest challenges in order to overcome issues with standardisation, availability in case of emergency, consumer engagement, as well as cost and crime reduction in the field of healthcare and finance. The application of DLT for personal DP is an emerging field with many challenges, as well as potential, as researchers and organisations continue to acknowledge it as promising solution for healthcare, finance and many present-day issues. Although not perfect and far from a one-size-fits-all solution, DLTs still satisfy numerous requirements, making them considerably suitable for personal DP approaches.
